\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates}

### 2022-09-07 manuscript.Rmd
abstract

authornote: |
  Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com).
abstract: |
  Visual illusions strikingly highlight how the brain uses contextual and prior information to inform our perception of reality. Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, which ideally require a controlled and gradual modulation of the effects of interest. In this set of studies, we used the parametric framework for visual illusions implemented in the *Pyllusion* software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength. We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task. We then extracted the participant-level performance scores (n=250), and provide evidence in favour of the existence of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions. Finally, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Negative Affect and Disinhibition. All the materials are available in open-access (https://github.com/RealityBending/IllusionGameValidation). We invite researchers to re-analyze the data using alternative approaches to provide complimentary findings on the effect, structure and correlates, of illusion sensitivity.

>>>>

authornote: |
  Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com).
abstract: |
  Visual illusions strikingly highlight how the brain uses contextual and prior information to inform our perception of reality. Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, which ideally require a controlled and gradual modulation of the effects of interest. In this set of studies, we used the parametric framework for visual illusions implemented in the *Pyllusion* software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength. We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250). Our results provide evidence in favour of the existence of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions. Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect. All the materials are available in open-access (https://github.com/RealityBending/IllusionGameValidation). We invite researchers to re-analyze the data using alternative approaches to provide complimentary findings on the effect, structure and correlates, of illusion sensitivity.
